---
title: "First Delivery Statistics"
author: "Eduardo Alarcón & Alfonso Pineda"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
markdown: 
    wrap: 82
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.Introduction:

The first delivery of the final project of Statistics of the degree of
Computer Science. In this document, we (Alfonso Pineda and Eduardo
Alarcón) will be showing the histogram of the main variable we have
chosen for our project, namely percentWeeksOnChart, which represents the
% of weeks with respect to a year the song has been on the top chart. As
well, we will have a Box Plot and the statistical Measures on the same
block as the histogram, the first one.

On the second block, we will show the "loudness" which we think is the
one that has the most relation with the main variable. We are also going
to show a Histogram, a Box Plot and the Statistical Measurements.

Lastly, we will show the Scatter Plot and the Linear Model

\newpage

# 1st Block:

Including the data from the excel: The first thing we need to do is
import the data we are going to work with.

```{r}
library(readxl)
SpotifySongs <- read_excel("spotistats.xlsx")
View(SpotifySongs)
```

### Histogram Then, we need to create the histogram, using R

```{r, out.height='70%'}
hist(SpotifySongs$percentWeeksOnChart, xlab = "% Of the year on Top charts",
     main = "% Of weeks a song is on the charts with respect to a year")

```

We have also created the histogram differentiating if the artist is a
Male, Female, Band or Collaboration between different artists (we used
different colors to view them):

```{r, fig.height = 4, fig.width = 6}

suppressWarnings(library(ggplot2))
SpotifySongs$people <- "Male"
SpotifySongs$people[SpotifySongs$GenderGroup == "F"] <- "Female"
SpotifySongs$people[SpotifySongs$GenderGroup == "Band"] <- "Band"
SpotifySongs$people[SpotifySongs$GenderGroup == "Colab"] <- "Colaboration"

qplot(percentWeeksOnChart, data=SpotifySongs, geom=c("histogram"), fill=people, 
      alpha=I(.65), main="Distribution people in top chartsr", 
      xlab="percentWeeksOnChart", ylab="Frequency")

```

## Boxplot

Then, we also need to create a Box Plot:

```{r, fig.height = 4, fig.width = 5}
boxplot(SpotifySongs$percentWeeksOnChart, ylab = "% Of weeks on top charts with respect to a year", 
        main = "Boxplot")
```

As we can see from the Box Plot there are some extreme outliers and
indicate that some of the songs stay on the top charts for significant
more time than the rest of songs.

We can see from the Box Plot that the Histogram is not symmetric at all.

# Statistical Measures

Now, it's time for us to calculate the statistical measures of the main
variable, percentWeeksOnChart. These measures are: the mean, the median,
the mode, the percentiles, the range, the variance, and the standard
deviation.

First, we need to store the variable as Data in R, then, we ask R to
describe the variable, which outputs the measures we need, as well as
the number of elements there are, in this case N: 100

```{r message=FALSE, warning=FALSE}
weeksonchart<-SpotifySongs$percentWeeksOnChart
suppressWarnings(library(summarytools))
# Describe the variable weeksonchart
descr(weeksonchart)
```

\newpage

# 2nd Block:

On this second part we will test which of the variables we have on our
study has the best correlation with the main variable. To asses this, we
will use the next block of R:

```{r}
# Choose best second variable
# Correlation between percentWeeksOnChart and Danceability\n"
cor(SpotifySongs$percentWeeksOnChart, SpotifySongs$Danceability)
# Correlation between percentWeeksOnChart and Duration\n"
cor(SpotifySongs$percentWeeksOnChart, SpotifySongs$Duration)
# Correlation between percentWeeksOnChart and Loudness\n"
cor(SpotifySongs$percentWeeksOnChart, SpotifySongs$Loudness)
```

With the previous results, we choose the variable Loudness. Here are the
statistical variables of the Loudness, as well as the Histogram and Box
Plot:

```{r}
loudness<-SpotifySongs$Loudness
descr(loudness)
```

```{r}
# Histogram/Box-Plot of the secondary variable
hist(loudness)
boxplot(loudness, horizontal = TRUE)
```

As we can see from the box plot and the histogram, this distribution is
more symmetrical than our main variable.

\newpage

# 3rd Block:

On the last block, we will see the Scatter Plot and Linear Model between
the main variable, the percentWeeksOnChart and the loudness.

```{r}
# Scatter plot without linear model of percentWeeksOnChart and Loudness
plot(
    SpotifySongs$Loudness, # We could also use loudness
    SpotifySongs$percentWeeksOnChart, 
    xlab = "Loudness",
    ylab = "Percentage weeks",
    col.lab = "gray19",
    col="skyblue3"
)
```
\newpage

## Scatter Plot with the Linear Model:

The scatter plot created with the Loudness is:

```{r}

RegressionModel <- lm(percentWeeksOnChart ~ Loudness, data=SpotifySongs)
print(RegressionModel)

plot(
    SpotifySongs$Loudness,
    SpotifySongs$percentWeeksOnChart,
    xlab = "Loudness",
    ylab = "Percentage weeks",
    col="skyblue3"
)
abline(RegressionModel, col="tomato3")


```


As we have not been able to create a valid regression model, 

We have to choose another variabel, as there is no sufficient correlation between them. 

```{r}
library(readxl)

#Import data
datos_statistics <- read_excel("spotistats.xlsx")

#Choose best second variable
cat("Correlation between percentWeeksOnChart and Danceability\n")
cat(cor(datos_statistics$percentWeeksOnChart, datos_statistics$Danceability))

cat("\n\nCorrelation between percentWeeksOnChart and Duration\n")
cat(cor(datos_statistics$percentWeeksOnChart, datos_statistics$Duration))

cat("\n\nCorrelation between percentWeeksOnChart and Loudness\n")
cat(cor(datos_statistics$percentWeeksOnChart, datos_statistics$Loudness))

cor()

#Histogram/Box-Plot of main variable
hist(datos_statistics$percentWeeksOnChart)
boxplot(datos_statistics$percentWeeksOnChart, horizontal = TRUE)

#Scatterplot without linear model of percentWeeksOnChart and Loudness
plot(
    datos_statistics$Loudness,
    datos_statistics$percentWeeksOnChart,
    xlab = "Loudness",
    ylab = "Percentage weeks"
)
```
